---
title: "Coronavirus Analysis in the United States"
knit: (function(input_file, encoding) {
  out_dir <- '/';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
author: "Non Nonceeya, Nicholas Guan, Angela Pandit"
date: "5/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## 1. Introduction

In the last couple months, the United States and many countries around the world have locked down and heavily restricted social gatherings in efforts to prevent the spread of SARS-CoV2, more commonly referred to as COVID-19. With the world essentially in lock down, it seems as though every news outlet is reporting about COVID-19. However, with the increasing articles about the virus and its effects, the amount of false information is increasing as well. 

https://www.worldometers.info/coronavirus/coronavirus-death-rate/

## 2. Motivation

Due to all the varying sources and data online, we decided to use a public dataset on COVID-19 cases within the United States to find out for ourselves what the true impacts of the virus are. Finding out the true impacts of COVID-19 is important because the virus is highly contagious. Therefore, the impacts of the virus will be widespread. Additionally, this tutorial will analyze the mortality rates within the United States by state without any media bias, so the results will show a more accurate picture of what the United States can expect in the future. 
In this tutorial, we will analyze the spread of coronavirus within the United States. We will examine the 

## 3, Libraries

As we continue with this analysis, we want to take advantage of many powerful R libraries.

```{r libs}
library(tidyr)
library(ggplot2)
library(broom)
library(caret)
library(tidyverse)
library(lubridate)
library(readr)
library(stringr)
```

## 4. Getting started with the Data

Dataset can be found at: https://www.kaggle.com/ashudata/covid19dataset 

There are **2** datasets available but for this analysis we have chosen the **COVID_Data_Basic** dataset

The data is fetched from John Hopkins University & European Centre for Disease Prevention and Control (ECDC), processed, and available in a CSV format. Also note that these sources are considered to be credible.

List of attributes in this dataset: 

* Country: Affected Country
* Date: Date of observation (YYYY-MM-DD)
* Confirmed: Cumulative # of confirmed cases
* Death: Cumulative # of death cases
* Recovered: Cumulative # of recovered cases
* newConfirmed: # of new confirmed cases/day
* newDeath: # of new death cases/day
* newRecovered: # of new recovered cases/day


First, let's take a look at the various rows that we have available in our dataset. As we can see, this dataset contains useful global Coronavirus data that we can use for analysis. Fortunately for us, the fields for our columns are appropiate for data analysis and we do not need to further **tidy** our data. Our countries are readily available as **Character vectors**, Date as **Dates**, and cases as **Doubles**. 

```{r data_setup}
covid_data <- read_csv("COVID_Data_Basic.csv")

sample_n(covid_data, 10)
```


Our goal is to analyze the COVID-19 situation in the United States so we would then use **filter** from the **dplyr** library to obtain just the United State data (labeled US).


```{r dataselect}

US_data <- filter(covid_data, Country == "US")

head(US_data, 10)

```

## 5. Exploratory Data Analysis

### Data Visualization

First, let's display a graph of the cumulative confirmed, death, and recovered cases to get a sense of the spread of the virus. In order to do this, we are going to take advantage of the powerful **ggplot** library. In addition to the cumulative cases, we will also graph the newConfirmed, newDeath, and newRecovered columns in another graph. The additional columns indicate the amount of cases on a given day and will assist us in further analysis. 

[Talk about central tendencies/skew]
```{r graph1}

US_data %>%
  ggplot(aes(Date)) +
  geom_line(aes(y=Confirmed, colour="Confirmed Cases (Cumulative)")) +
  geom_line(aes(y=Death, colour="Deaths (Cumulative)")) +
  geom_line(aes(y=Recovered, colour="Recovered (Cumulative)")) +
  labs(title="Cases in the United States", x="Month (2020)", y="Amount")

US_data %>%
  ggplot(aes(Date)) +
  geom_line(aes(y=newConfirmed, colour="Confirmed Cases")) +
  geom_line(aes(y=newDeath, colour="Deaths")) +
  geom_line(aes(y=newRecovered, colour="Recovered")) +
  labs(title="Cases per day in the U.S.", x="Month (2020)", y="Amount")
```


We can see that starting from mid-March, there was a very steep increase in cases per day. Let's filter the date to after March 15th using **filter** again so we can get a better view at the growth. In order to get a better picture after mid-March, we will once again graph the two respective graphs. 

```{r graph2}

US_data_filt <- filter(US_data, Date >= "2020-03-15")

US_data_filt %>%
  ggplot(aes(Date)) +
  geom_line(aes(y=Confirmed, colour="Confirmed Cases")) +
  geom_line(aes(y=Death, colour="Deaths")) +
  geom_line(aes(y=Recovered, colour="Recovered")) +
  labs(title="Cumulative Cases in U.S.", x="Month (2020)", y="Amount")

US_data_filt %>%
  ggplot(aes(Date)) +
  geom_line(aes(y=newConfirmed, colour="Confirmed Cases")) +
  geom_line(aes(y=newDeath, colour="Deaths")) +
  geom_line(aes(y=newRecovered, colour="Recovered")) +
  labs(title="Cases per day in the U.S.", x="Months (March 15... (2020))", y="Amount")

```


[throw in some data articles talking about spike, lack of materials]

[credible sources so we should be concerned and lessen this curve]

## 6. Hypothesis testing and Machine Learning 

[can essentially go off project5 code & use random forest to predict growth and use this as warning to stay home]

## 7. Conclusion
